---
title: "Logistic Regression for predicting "Time to Relapse" for Visceral Leishmaniasis individuals
author: "rwomersley"
date: "03/11/2021"
output: html_document
---
# Logistic Regression for Prediction of Relapse in Visceral Leishmaniasis
### Code used for performing a logistic regression to predict the time to relapse for patients with VL, based on levels of CD4 cells, PD1 and IFNg expression.

## Part 1
### Multinomial Logistic Regression to predict 'Time to Relapse' using CD4, PD1 and IFNg levels.
*Load in the data and neccessary packages*
```{r, echo=FALSE}
library("readxl")
library("nnet")
library("stargazer")
library("ggplot2")
library("pROC")


#load in data
setwd("") # set working directory - input your working directory here
Scores <- read_excel("Raw_Values_LogReg_Visceral_Leishmaniasis.xlsx", sheet = "Score", skip = 1)

#--------------------------------------------------------------------------------
#data cleaning and preprocessing
#relapse group
relapse_group <- Scores[,6:8]
relapse_tps <- Scores[,9] 
#remove missing values
relapse_group <- relapse_group[c(2:6,8:16, 18:21),]
relapse_tps <- relapse_tps[c(2:6,8:16, 18:21),]
#rename colnames 
colnames(relapse_group) <- c("CD4", "IFNg", "PD1")
colnames(relapse_tps) <- "Time to Relapse"
#add column - Infection and time to relapse
relapse_group$Infection <- as.factor("VLHIV")
relapse_group$Relapse <- 1
relapse_tps$`Time to Relapse` <- ifelse(relapse_tps$`Time to Relapse`=="12m", "Relapse at 6-12M", "Relapse at 3M")
relapse_tps$`Time to Relapse` <- as.factor(relapse_tps$`Time to Relapse`)
relapse_vlhiv <- cbind(relapse_group, relapse_tps)
#remove mistyped values (due to being out of threshold of quantification)
relapse_vlhiv <- relapse_vlhiv[-2,] 
#--------------------------------------------------------------------------------
#no relapse group
no_relapse <- Scores[,1:3]
no_relapse_tps <- Scores[, 4]
no_relapse$`CD4 score...1` <- as.numeric(no_relapse$`CD4 score...1`)
#remove missing values
no_relapse <- no_relapse[c(4:6, 12, 14:27),]
no_relapse_tps <- no_relapse_tps[c(4:6, 12, 14:27),]
#rename colnames 
colnames(no_relapse) <- c("CD4", "IFNg", "PD1")
colnames(no_relapse_tps) <- as.factor("Time to Relapse")
#add column - Infection and time to relapse
no_relapse$Infection <- as.factor("VLHIV")
no_relapse$Relapse <- 0
no_relapse_tps$`Time to Relapse` <- ifelse(no_relapse_tps$`Time to Relapse`=="relapsedafter 6-12 months and before 3 years3 year", "Relapse at 12M+", "No Relapse in Study")
no_relapse_tps$`Time to Relapse` <- as.factor(no_relapse_tps$`Time to Relapse`)
norelapse_vlhiv <- cbind(no_relapse, no_relapse_tps)
#remove extreme values (out of capability of quantification)
norelapse_vlhiv <- norelapse_vlhiv[-7,] 




#put them all together! 
full <- rbind(relapse_vlhiv, norelapse_vlhiv)

#prep data, we want three outcomes now for relapse
full$multi <- full$`Time to Relapse`
full$multi[full$multi == "Relapse at 12M+"] <- "No Relapse in Study"
full$multi <- droplevels(full$multi)
full$multi <- factor(full$multi, levels = c("No Relapse in Study", "Relapse at 6-12M", "Relapse at 3M"))

multi <- full[,c(1:3, 7)]

```

#Build Model and evaluate coefficients

```{r}
multi_model <- multinom(multi ~ CD4 * IFNg * PD1, data = multi)


summary(multi_model)
stargazer(multi_model, type="text")

```

##predict class
```{r}
# predictions
#class
pred <- predict(multi_model, multi)

#predicted probabilities?
pp <- fitted(multi_model)

#make a new dataframe with prediction probabilities. 
prediction <- as.data.frame(cbind(pp, pp))

prediction[,4] <- ifelse(prediction[,4]>=0.5, 1, 0)
prediction[,5] <- ifelse(prediction[,5]>=0.5, 1, 0)
prediction[,6] <- ifelse(prediction[,6]>=0.5, 1, 0)
prediction$actual <- multi$multi
colnames(prediction) <- c("Probability - No Relapse", "Probability - Relapse at 6-12M", "Probability - Relapse at 3M", "Predict - No Relapse", "Predict - Relapse at 6-12M", "Predict - Relapse at 3M", "Correct Classification")

colours <- c("Relapse at 3M" = "#FF0000", "Relapse at 6-12M" = "#FF8000", "No Relapse in Study" = "#0000FF")

#No relapse ins study
ggplot(prediction, aes(x = `Correct Classification`, y = `Probability - No Relapse`)) + geom_boxplot( linetype = "dotted") +
  geom_boxplot(aes(ymin=..lower.., ymax=..upper..), fill = colours) +
  geom_jitter(shape=16, position=position_jitter(0.2)) +  theme_bw()
```


Probability of relapsing between 6-12m
```{r}
#6-12m
ggplot(prediction, aes(x = `Correct Classification`, y = `Probability - Relapse at 6-12M`)) + geom_boxplot( linetype = "dotted") +

  geom_boxplot(aes(ymin=..lower.., ymax=..upper..), fill = colours) +
  geom_jitter(shape=16, position=position_jitter(0.2)) +  theme_bw()
```


Probability of relapsing at 3m
```{r}
#3m
ggplot(prediction, aes(x = `Correct Classification`, y = `Probability - Relapse at 3M`)) + geom_boxplot( linetype = "dotted") +
  geom_boxplot(aes(ymin=..lower.., ymax=..upper..), fill = colours) +
  geom_jitter(shape=16, position=position_jitter(0.2)) +  theme_bw()
```

*Receiver Operating Curve (ROC) for results, and calculating Area Under ROC (AUROC)*
```{r}
prediction$Correct_3M <- ifelse(prediction$`Correct Classification`=='Relapse at 3M', 1,0)
prediction$Correct_612m <- ifelse(prediction$'Correct Classification' == 'Relapse at 6-12M', 1, 0)
prediction$Correct_Norelapse <- ifelse(prediction$'Correct Classification' == 'No Relapse in Study', 1, 0)

#based on probability
roc <- roc(response = prediction$Correct_3M, predictor = prediction$`Probability - Relapse at 3M`, 
                plot=TRUE, legacy.axes=TRUE, col="#606060", lwd=4 , print.auc=TRUE, print.auc.x=20, 
                ci=TRUE, ci.type="bars", 
                smooth = TRUE,
                percent = TRUE)
plot.roc(response = prediction$Correct_612m, prediction$`Probability - Relapse at 6-12M`, 
         col="#C0C0C0", lwd=4, print.auc=TRUE, print.auc.x=20, print.auc.y=45, 
         smooth = TRUE, add = TRUE,
         ci=TRUE, percent = TRUE)
legend("bottomright", legend=c("Relapse at 3M", "Relapse at 6-12M"), col=c("#606060", "#C0C0C0"), lwd=4)
```













